---
author: "Lilly McClendon"
date: "2025-02-07"
title: "Portfolio Piece 01"
output: github_document
---

#### The purpose of Portfolio Piece 1 is to create plots of squirrel sightings in New York City and pizza locations in New York City. Using that data, use latitutde and longitude from both databases to see if there is a visual connection between squirrel sightings and pizza shop locations. Then the data will be transformed.

### Data Credits 

[Data from NYC Squirrel Census](https://www.thesquirrelcensus.com/). Retrieved from [jonthegeek](https://github.com/rfordatascience/tidytuesday/tree/main/data/2019/2019-10-29) originally shared by Sara Stoudt. [Pizza Data](https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-01/pizza_datafiniti.csv")

### Load squirrel data and packages 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r get-squirrel-data}
nyc_squirrels <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2019/2019-10-29/nyc_squirrels.csv")
```
### Squirrel fur color visualization 

```{r squirrel-color}
ggplot(data=nyc_squirrels, aes(x=primary_fur_color, fill = primary_fur_color)) + 
  geom_bar(stat = "count", width = 1) + 
  scale_fill_manual(values = c("black", "chocolate4","darkgrey")) + 
  labs(title="NYC Squirrel Fur Color", 
       xlab = "Number of Observations", 
       ylab = "Fur Color")
```

#### This visualization utilizes the scale_fill_manual function in order to customize the colors of the bars to match the fur color they represent. 

### Load pizza data 

```{r load-pizza-data}
pizza_datafiniti <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-01/pizza_datafiniti.csv")
```

```{r filter-NY-only}
pizza_nyc <- pizza_datafiniti %>%
  filter(city == "New York")
```

### Combine squirrel and pizza data 
