---
date: "2025-05-01"
title: "Portfolio 10"
output:
  html_document:
    toc: true
    toc_float: true
---

The purpose of the last portfolio piece is to create a document that has code chunks for various tasks that can be easily accessed and referred to. 

# Codes {.tabset}

## Data processing 

### Filter data 

```{r filter}
# ```{r filter}
# filtered_data <- data %>%
#  filter(dataset == "column_want_to_filter_for")
# ```
```

### Filter data for column value of certain value 

```{r column_value_over_number}
# ```{r column_value_over_number}
# data_name %>%
#  filter(column_name > 1.0)
# ```
```


### Create new column 

```{r mutate, eval=TRUE}
# ```{r mutate, eval=TRUE}
# new_data_name <- data_name %>% 
#  mutate( 
#    column_name = if_else(born_country == "value1", "value2", "value3"))
# ```
```


### Join datasets 

```{r join_data}
# ```{r join_data}
# data_combined <- full_join(data1, data2, by = "shared_column_name")
# data_combined
# ```
```


### Variable Names 

```{r variable_names}
# ```{r variable_names}
# library(dplyr)
# count(data_name)
# colnames(data_name)
# ```
```


### Separate values (uoe example)

```{r separate-title_uoe_example, error = TRUE}
# ```{r separate-title_uoe_example, error = TRUE}
# uoe_art <- uoe_art %>%
#  separate(title, into = c("title", "date"), sep = "\\(") %>%
#  mutate(year = str_remove(date, "\\)") %>% as.numeric()) %>%
#  select(title, artist, year, date)
# ```
```


## Charts 

### Scatter plot 

```{r scatter_plot, fig.height=3, fig.width=6}
# ```{r scatter_plot, fig.height=3, fig.width=6}
# ggplot(data = your_data_name, mapping = aes(x = x, y = y)) +
# geom_point()
# ```
```


### Scatter plot with colored dots 

```{r scatter_plot_colors}
# ```{r scatter_plot_colors}
# ggplot(data_name, aes(x = column_name, y = column_name, color = column_name)) + 
#  geom_point() + 
#  labs(title = "Figure Title")
# ```
```


### Histogram 

```{r histogram}
# ```{r histogram}
# ggplot(data = data_name, aes(x = column_name)) +
#  geom_histogram(binwidth = 0.2) + 
#  labs(title = "Figure Title")
# ```
```


### Histogram with facet 

```{r histogram-facet-by-for_example_country}
# Will produce multiple histograms 
# ```{r histogram-facet-by-for_example_country}
# ggplot(plastic_waste, aes(x=column_name)) +
#  geom_histogram() +
#  facet_wrap(~continent) +
#  labs(title = "Column Name (e.g., Plastic waste per capita) by country")
# ```
```


### Bar plot with facet 

```{r faceted_bar_plot}
# ```{r faceted_bar_plot}
#data_name %>% 
#  ggplot(aes(x=column_name)) + 
#  geom_bar() + 
#  ggtitle("Figure Title") + 
#  facet_wrap(~column_name)
# ``` 
```


### Horizontal Bar plot with facet 

```{r horizontal_faceted_bar_plot}
# ```{r horizontal_faceted_bar_plot}
# data_name %>% 
#  ggplot(aes(x=column_name)) + 
#  geom_bar() + 
#  ggtitle("Figure Title") + 
#  facet_wrap(~column_name) + 
#  coord_flip()
# ``` 
```


### Stacked bar plot 

```{r plot-bornus-resideus, eval = TRUE}
# ```{r plot-bornus-resideus, eval = TRUE}
# ggplot(data_name, aes(x=column_name, fill=column_name, color=same_as_fill)) + 
#  geom_bar(stat = "count") +
#  ggtitle("Figure Title") +
#  facet_wrap(~column)name
# ```
```


### Density plot 

```{r density_plot}
# ```{r density_plot}
# ggplot(data = data_name, aes(x = column_name)) +
#  geom_density() + 
#  labs(title = "Figure Title")
# ```
```


### Overlaid density plot 

```{r density_plot_multiple_lines}
# ```{r density_plot_multiple_lines}
# ggplot(data = data_name, aes(x = column_name, shape = column_name_for_variable_want_each_line_to_represent)) +
#  geom_density() + 
#  labs(title = "Figure Title")
# ```
```


### Color-coded overlaid Density Plot 

```{r desnity_plot_multiple_lines_color_coded}
# ```{r desnity_plot_multiple_lines_color_coded}
# ggplot(data = data_name, mapping = aes(x = column_name, color = column_name_for_variable_want_each_line_to_represent)) +
# geom_density() + 
#  labs(title = "Figure Title")
# ```
```


### Density plot filled curves 

```{r density_curve_fill_colored}
# ```{r density_curve_fill_colored}
# ggplot(data = data_name, mapping = aes(x = column_name, color = column_name_for_variable_want_each_line_to_represent, fill = same_column_name_as_color)) +
#  geom_density() + 
#  labs(title = "Figure Title")
# ``` 
```


### Density plot filled curves transparency 

```{r density_curve_fill_colored_transparency}
# ```{r density_curve_fill_colored_transparency}
# ggplot(data = data_name, mapping = aes(x = column_name, color = column_name_for_variable_want_each_line_to_represent, fill = same_column_name_as_color)) +
# geom_density(alpha = .25) + 
#  labs(title = "Figure Title")
# ```
```


### Side by side box plots 

```{r side_by_side_box_plots}
# ```{r side_by_side_box_plots}
# ggplot(data = data_name, mapping = aes(x = column_name, y = column_name)) +
#  geom_boxplot() + 
#  labs(title = "Figure Title")
# ```
```


### Violin plots 

```{r violin_plot}
# ```{r violin_plot}
# ggplot(data_name, aes(x = column_name, y=column_name))+
#  geom_violin() + 
#  labs(title = "Figure Title")
# ```
```


### Scatter plot with longitude and latitude 

```{r scatter_plot_lat_long}
# ```{r scatter_plot_lat_long}
# ggplot(dataPname) + 
#  geom_point(mapping = aes(x = longitude, y = latitude,
#  color = column_name)) + 
#  labs(
#    title= "Figure Title",
#    x = "Longitude", 
#    y = "Latitude",
#    tag = "Figure 1")
# ```
```


### Line plot 

```{r line-plot}
# ```{r line-plot}
# library(ggplot2)
# data_name %>%
#  ggplot(aes(
#    x = column_name,
#    y = column_name,
#    group = column_name,
#    color = same_as_group)) + 
#  labs(title = "Figure Title", x = "X axis label", y = "Y axis label") + 
#  geom_line() + 
#  guides(color = guide_legend(title = "Lengend Title"))
# ``` 
```


### Bar plot descending and x axis label tilted 

```{r bar_plot_desc_xlabel_tilt}
# ```{r bar_plot_desc_xlabel_tilt}
# library(ggplot2)
# ggplot(data_name, aes(x=column_name)) + 
#  geom_bar(aes(x=forcats::fct_infreq(column_name))) + 
#  theme(axis.text.x = element_text(angle = 20, vjust = .5))
# ```
```


### Arrange bar plot in specific order of x axis label (age group example)

```{r demographic_distribution_age_cat_reorder}
# ```{r demographic_distribution_age_cat_reorder}
# library(ggplot2)
# library(tidyverse)

# agecat_compas <- subset(compas, select=c("age_cat"))

# agecat_compas$age_cat <- factor(agecat_compas$age_cat, levels = c ("Less than 25", "25 - 45", "Greater than 45"))

# ggplot(agecat_compas, aes(x=age_cat)) + 
#  geom_bar() + 
#  labs(x="Age Category", y = "Count")
# ```
```


### Jitter plot 

```{r jitterplot}
# ```{r jitterplot}
# library(ggplot2)
# ggplot(data_name, aes(x=column_name, y=column_name)) + 
#  geom_jitter()
# ``` 
```


### Jitter plot with regression line example 

```{r jitterplot_regression_line}
# ```{r jitterplot_regression_line}
# library(ggplot2)
# ggplot(evals, aes(x=bty_avg, y=score)) + 
#  geom_jitter() +
#  geom_smooth(method='lm', color="orange", se=FALSE)
# ```
```


## Calculations 

### Correlation calculation 

```{r correlation}
# ```{r correlation}
# your_data_name %>%
#   summarize(r = cor(x, y))
# ```
```


### Number of rows 
```{r number_rows}
# ```{r number_rows}
# new_data <- original_data %>%
#  filter(column_name == "value_filter_for")
# nrow(new_data)
# view(new_data)
# ```
```


### Count 

```{r count}
# ```{r count}
# library(dplyr)
# data_name %>% 
# count(column_name)%>%
# arrange(desc(n))
# ``` 
```


### Count regardless of capitalization (uoe example)

```{r child_in_title}
### ```{r child_in_title}
### library(stringr)
### library(dplyr)

### count_title_includes_child <- sum(str_detect(uoe_art_corrected$title, regex("child\\b", ignore_case = TRUE)))

### count_title_includes_children <- sum(str_detect(uoe_art_corrected$title, regex("child", ignore_case = TRUE)))

### print(count_title_includes_child)
### print(count_title_includes_children)
# ``` 
```


### Descriptive Statistics 

``` {r descriptive-stats}
# ``` {r descriptive-stats}
# summary(data_name)
# IQR(data_name$closest)
# table <- table(data_name)
# print(table)
# ```
```


### Mean and SD

```{r mean_sd}
# ```{r mean_sd}
# summarize(data_name, 
#          mean_score = mean(score_or_variable_name), 
#         sd_score = sd(score_or_variable_name))
# ``` 
```


### IQR 

```{r IQR}
# ```{r IQR}
# IQR(data_name$closest)
# IQR_data_name <- [insert numerical IQR value here] 
# quart1_data_name <- [insert numerical IQR value here] 
# quart3_data_name <- [insert numerical IQR value here] 
# print(quart1_data_name - (IQR_data_name*1.5))
# print(quart3_data_name + (IQR_data_name*1.5))
# ```
```


### Combinations 

```{r combinations}
# ```{r combinations}
# total_combinations <- nrow(data1) * nrow(data2)
# print(total_combinations)
# ```
```


### Haversine formula 

```{r Haversine-distance-formula}
# ```{r Haversine-distance-formula}
# haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
#  long1 <- long1 * pi / 180
#  lat1 <- lat1 * pi / 180
#  long2 <- long2 * pi / 180
#  lat2 <- lat2 * pi / 180

#  R <- 6371 # Earth mean radius in km

#  a <- sin((lat2 - lat1) / 2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1) / 2)^2
#  d <- R * 2 * asin(sqrt(a))

#  return(round(d, round)) # distance in km
# }
# ```
```


### Calculate distance and sort by minimum distance 

```{r add_distance}
# ```{r add_distance}
# data_name <- data_name %>%
#  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))

# data_name_mindist <- data_name %>%
#  group_by(address.x) %>%
#  summarize(closest = min(distance))
# ```
```


## Regression Formulas Examples 

### Linear regression with numerical predictor example 

```{r linear_reg_fit}
# ```{r linear_reg_fit}
# m_bty <- linear_reg() %>% 
# set_engine("lm") %>% 
#  fit(score ~ bty_avg, data = evals) %>% 
# tidy()
# ``` 
```

$\hat{\text{Score}}_i$ = 3.88 + 0.07 x bty_avg$_i$


### R-squared example 

```{r m_bty-rsquared}
# ```{r m_bty-rsquared}
# library(tidymodels)
# library(broom)

# R_m_bty <- linear_reg() %>% 
#  set_engine("lm") %>% 
#  fit(score ~ bty_avg, data = evals)

# broom::glance(R_m_bty$fit)
# ``` 
```


### Linear regression with categorical predictor example 

```{r regression_gender_scores}
# ```{r regression_gender_scores}
# m_gen <- lm(score ~ gender, data = evals)
# tidy(m_gen)
# contrasts(evals$gender)
# ```
```


### Regression with new reference group example

```{r m_rank_relevel_model}
# ```{r m_rank_relevel_model}
# m_rank_relevel <- lm(score ~ rank_relevel, data = evals)
# tidy(m_rank_relevel)
# contrasts(evals$rank_relevel)

# library(tidymodels)
# library(broom)

# R_m_rank_relevel <- linear_reg() %>% 
#  set_engine("lm") %>% 
#  fit(score ~ rank_relevel, data = evals)

# broom::glance(R_m_rank_relevel$fit)
# ``` 
```


### Multiple linear regression example 

```{r linear-regression-bty_avg-gender}
# ```{r linear-regression-bty_avg-gender}
# m_bty_gen <- linear_reg() %>% 
#  set_engine("lm") %>% 
#  fit(score ~ bty_avg + gender, data = evals)
# tidy(m_bty_gen)
# contrasts(evals$gender)

# broom::glance(m_bty_gen$fit)

# tidy(m_bty_gen) %>% 
# mutate(exp_estimate = exp(estimate)) %>% 
#  select(term, estimate, exp_estimate)
# ``` 
```


### Search for best model (backward selection) example 

```{r backward-selection}
# ```{r backward-selection}
# library(MASS)

# lm_full_model <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval + cls_students + cls_level + cls_profs + cls_credits + bty_avg, data = evals)

# best_model <- stepAIC(lm_full_model, direction = "backward", trace = 1)
# summary(best_model)

# best_model_only_best <- stepAIC(lm_full_model, direction = "backward", trace = 0)
# summary(best_model_only_best)
# ```
```


### Full model of best model example 

```{r linear-regression-full_model-best-fit}
# ```{r linear-regression-full_model-best-fit}
# m_best_full_model <- linear_reg() %>% 
#  set_engine("lm") %>% 
#  fit(score ~  ethnicity + gender + cls_perc_eval + cls_credits + bty_avg, data = evals) %>% 
# tidy()
# view(m_best_full_model)

# library(tidymodels)
# library(broom)

# R_m_best_full_model <- linear_reg() %>% 
#  set_engine("lm") %>% 
#  fit(score ~ ethnicity + gender + cls_perc_eval + cls_credits + bty_avg, data = evals)

# broom::glance(R_m_best_full_model$fit)

# contrasts(evals$gender)
# contrasts(evals$ethnicity)
# contrasts(evals$cls_credits)
# ``` 
```


## Emojis 

### Emojis 

```{r emojis}
# ```{r emojis}
# install.packages("emoji")
# library(emojifont)
# search_emoji('smile')
# cat("Text to be displayed", emoji("smiley_cat"))
# ```
```

